---
title: "Open knowledge in R with Wikimedia APIs"
author: "Mikhail Popov"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  beamer_presentation:
    highlight: kate
    includes:
      in_header: header.tex
    keep_tex: yes
    latex_engine: xelatex
    md_extensions: -autolink_bare_uris+hard_line_breaks+startnum+definition_lists+footnotes
  md_document:
    includes:
      before_body: prefix.md
    toc: yes
    toc_depth: 2
    variant: markdown_github+backtick_code_blocks+autolink_bare_uris+startnum+definition_lists+footnotes
institute: Wikimedia Foundation
---

## Introduction

[Wikimedia Foundation](https://wikimediafoundation.org/) is a non-profit that operates free & open projects like [Wikipedia](https://www.wikipedia.org/), [Wiktionary](https://www.wiktionary.org/), and [Wikidata](https://www.wikidata.org/) that anyone can contribute to

No time to talk about me (plus that's always the boring part)[^me]

[^me]: If you're **really** curious just search for [User:MPopov (WMF)](https://meta.wikimedia.org/wiki/User:MPopov_(WMF)) on [Meta-Wiki](https://meta.wikimedia.org/)

A Markdown copy of this deck is at [git.io/vSi6a](https://git.io/vSi6a) for following along

R packages required to follow along:

```{r install, eval=FALSE}
install.packages(
  c("pageviews", "WikipediR", "WikidataR",
    "WikidataQueryServiceR", "magrittr"),
  repos = c(CRAN = "https://cran.rstudio.com")
)
```

## Session Info

```{r session_info, echo=FALSE}
library(magrittr)
session_info <- sessionInfo(package = c(
  "pageviews", "WikipediR", "WikidataR",
  "WikidataQueryServiceR"
))
pkgs <- dplyr::bind_rows(session_info$otherPkgs)
pkgs$Imports <- vapply(regmatches(pkgs$Imports, gregexpr("([a-zA-Z]+)", pkgs$Imports)), paste, "", collapse = ", ")
```

- Running R `r getRversion()` on `r session_info$running`
- Rendered with [rmarkdown](http://rmarkdown.rstudio.com/) `r packageVersion("rmarkdown")` and [knitr](https://yihui.name/knitr/) `r packageVersion("knitr")`
- The pipe (`%>%`) from [magrittr](https://cran.r-project.org/package=magrittr) is **occasionally** used
- Using the following versions of packages for demos:

```{r pkgs, echo=FALSE}
knitr::kable(pkgs[, c("Package", "Version", "Imports")])
```

## Pageviews

WMF provides an [API for accessing daily and monthly pageviews of any article on any project](https://wikitech.wikimedia.org/wiki/Analytics/PageviewAPI) for counts from 2015 onwards.[^old] The package [pageviews](https://cran.r-project.org/package=pageviews) allows you to get those counts in R:

[^old]: [wikipediatrend](https://cran.r-project.org/package=wikipediatrend) package wraps the [stats.grok.se](http://stats.grok.se/) API which has historical Wikipedia pageview data for 2008 up to 2016 from [these pageview count dumps](https://dumps.wikimedia.org/other/pagecounts-raw/).

```{r pageviews_data, cache=TRUE}
library(pageviews)
r_pageviews <- article_pageviews(
  project = "en.wikipedia",
  article = "R (programming language)",
  user_type = "user", start = "2015100100",
  end = format(Sys.time(), "%Y%m%d%H%M00")
)
```

----

```{r pageviews_visualization, echo=knitr::opts_knit$get("rmarkdown.pandoc.to") != "beamer", dev='png', dpi=300, fig.align='center', out.width='\\linewidth', fig.path='figures/'}
library(ggplot2)
r_pageviews$date %<>% as.Date()
ggplot(r_pageviews, aes(x = date, y = views)) +
  geom_line(color = rgb(0, 102, 153, maxColorValue = 255)) +
  geom_text(data = dplyr::top_n(r_pageviews, 1, views),
            aes(x = date, y = views, label = format(date, "%d %B %Y"),
                hjust = "left"), nudge_x = 10, size = 6) +
  geom_point(data = dplyr::top_n(r_pageviews, 1, views),
             aes(x = date, y = views), color = rgb(153/255, 0, 0)) +
  scale_y_continuous(
    breaks = seq(2e3, 10e3, 1e3),
    labels = function(x) { return(sprintf("%.0fK", x/1e3)) }
  ) +
  scale_x_date(date_breaks = "2 months", date_labels = "%b\n%Y") +
  labs(x = NULL, y = "Pageviews",
       title = "Daily pageviews of R's entry on English Wikipedia",
       subtitle = "Desktop and mobile traffic, excluding known bots") +
  theme_minimal(base_size = 18, base_family = "Gill Sans")
```

## Wikidata

- [Wikidata](https://www.wikidata.org/) is a language-agnostic open knowledge base
- Facts are expressed as 3-part statements:
    - Subject (resource)
    - Predicate (property type)
    - Object (property value, can be another resource)
- Examples:
    - "R" ([Q206904](https://www.wikidata.org/wiki/Q206904)) is an "instance of" ([P31](https://www.wikidata.org/wiki/Property:P31)) a "programming language" ([Q9143](https://www.wikidata.org/wiki/Q9143))
    - "RStudio" ([Q4798119](https://www.wikidata.org/wiki/Q4798119)) was "programmed in" ([P277](https://www.wikidata.org/wiki/Property:P277)) "C++" ([Q2407](https://www.wikidata.org/wiki/Q2407))
    - "Portland" ([Q6106](https://www.wikidata.org/wiki/Q6106)) had a "population" ([P1082](https://www.wikidata.org/wiki/Property:P1082)) of 583,776 (in 2010)
- Resources and properties have unique numeric identifiers but can have human-friendly labels in any language

## Wikidata Query Service (WDQS)

- Allows querying Wikidata with [SPARQL](https://en.wikipedia.org/wiki/SPARQL)
- Provides a public SPARQL endpoint usable via:
    - Web front-end: [query.wikidata.org](https://query.wikidata.org/)
    - Web API (`https://query.wikidata.org/sparql?query=<SPARQL>`)
    - In Python with [SPARQLWrapper](https://rdflib.github.io/sparqlwrapper/)
    - In R with:
        - [SPARQL](https://cran.r-project.org/package=SPARQL) package
        - [WikidataQueryServiceR](https://cran.r-project.org/package=WikidataQueryServiceR)
- For useful reference links, see
  `help("WDQS", package = "WikidataQueryServiceR")`

## Basic SPARQL Example

```
# PREFIXes are optional when using WDQS
PREFIX wd: <http://www.wikidata.org/entity/>
PREFIX wdt: <http://www.wikidata.org/prop/direct/>
PREFIX wikibase: <http://wikiba.se/ontology#>
PREFIX bd: <http://www.bigdata.com/rdf#>

SELECT DISTINCT ?instanceOfLabel
WHERE {
  wd:Q206904 wdt:P31 ?instanceOf .
  SERVICE wikibase:label {
    bd:serviceParam wikibase:language "en"
  }
}
```

----

```{r wdqs_basic, cache=TRUE, message=FALSE}
library(WikidataQueryServiceR)
query_wikidata('SELECT DISTINCT ?instanceOfLabel
WHERE {
  wd:Q206904 wdt:P31 ?instanceOf .
  SERVICE wikibase:label {
    bd:serviceParam wikibase:language "en"
  }
}') %>% head(n = 5L)
```

## Advanced SPARQL Example

- Prefix `wd:` points to an entity
- Prefix `p:` points not to the object, but to a statement node
- Prefix `ps:` within the statement node retrieves the object (value)
- Prefix `pq:` within the statement node retrieves the qualifier info

```{r wdqs_advanced_query, cache=TRUE}
r_versions_query <- "SELECT DISTINCT
  ?softwareVersion ?publicationDate
WHERE {
  BIND(wd:Q206904 AS ?R)
  ?R p:P348 [
    ps:P348 ?softwareVersion;
    pq:P577 ?publicationDate
  ] .
}"
```

----

```{r wdqs_advanced_results, cache=TRUE, dependson='wdqs_advanced_query', message=FALSE}
r_versions_results <- query_wikidata(r_versions_query)
```

### Results

```{r wdqs_advanced_print, results='asis', echo=FALSE}
r_versions_results <- dplyr::arrange(r_versions_results, publicationDate)
knitr::kable(rbind(
  head(r_versions_results, 3),
  data.frame(softwareVersion = "...", publicationDate = "..."),
  tail(r_versions_results, 3)
), row.names = FALSE)
```

## Final Remarks

Source for the whole shebang is up on GitHub: [bearloga/wmf](https://github.com/bearloga/wmf/tree/master/presentations/talks/Cascadia%20R%20Conference%202017), available under [CC BY-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/)

Specifically: wmf/presentations/talks/Cascadia R Conference 2017/

### Contact Info

- **Twitter**: [bearloga](https://twitter.com/bearloga)
- **WMF-related**: [mikhail@wikimedia.org](mailto:mikhail@wikimedia.org)
  (PGP public key: [people.wikimedia.org/~bearloga/public.asc](https://people.wikimedia.org/~bearloga/public.asc))
- **General**:  [mikhail@mpopov.com](mailto:mikhail@mpopov.com)
  (PGP public key on [keybase.io/mikhailpopov](https://keybase.io/mikhailpopov))
